Randomly-Scoped Lambda Calculus

### 2012-12-19 RSLC/paper/paper.tex
write more

\usepackage{yfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathpartir}
\usepackage{url}
\usepackage{graphics}
\usepackage{marvosym}
\usepackage{stmaryrd}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utopia]{mathdesign}

@@

Anyway, Figure~\ref{fig:eval} shows the evaluation rules, which are basically what you'd expect, and Figure~\ref{fig:subst} shows the substitution rules, which are new.
In the normal $\lambda$-calculus, in the {\sc Subst-Capture} rule, substitution stops when a new binding's name is the same as the one being substituted for, but in RSLC we switch to a second substitution mode, {\sc Capture}, in which substituted expressions are appended to a variable's $\rep{e}$ list (in the {\sc Capture-Var} rule).
Similarly, when substituting for a variable before switching to {\sc Capture} mode (in the {\sc Subst-Var} rule), we ... hmm, why is two-moded substitution needed with DRS anyway?

@@

\begin{figure}[t]
	\fbox{$\subst{e_0}{x}{e}$}
	\TypeRules{
		\SubstLamCapture \qquad \SubstLam \\
		\SubstVarX \qquad \SubstVarY \\
	}
	\fbox{$\capture{e_0}{x}{e}$}
	\TypeRules{
		\CaptureVarX \qquad \CaptureVarY \\
	}
	\caption{RSLC substitution rules. The ``$e_1 \in ...$'' premise in {\sc Subst-Var-X} represents nondeterministic choice. (The complete set of rules is shown in Figure~\ref{fig:subst-complete}.)}
	\label{fig:subst}
\end{figure}

@@

\begin{figure*}[t]
	\fbox{$\subst{e_0}{x}{e}$}
	\TypeRules{
		\SubstLamCapture \qquad \SubstLam \qquad
		\SubstApp \\
		\SubstVarX \qquad \SubstVarY \\
		\SubstCaseCapture \\ \SubstCase \\
		\SubstZero \qquad \SubstSuc \\
		\SubstLetCapture \qquad \SubstLet
	}
	\fbox{$\capture{e_0}{x}{e}$}
	\TypeRules{
		\CaptureLam \qquad \CaptureApp \qquad
		\CaptureVarX \qquad \CaptureVarY \\
		\CaptureCase \\
		\CaptureZero \qquad \CaptureSuc \qquad \CaptureLet
	}
	\caption{Complete RSLC substitution and capture rules.}
	\label{fig:subst-complete}
\end{figure*}

>>>>

\usepackage{yfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathpartir}
\usepackage{url}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{marvosym}
\usepackage{stmaryrd}
\usepackage{epsdice}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utopia]{mathdesign}

@@

Anyway, Figure~\ref{fig:eval} shows the evaluation rules, which are basically what you'd expect, and Figure~\ref{fig:subst} shows the substitution rules, which are new.
In the normal $\lambda$-calculus, substitution would stop when a new binding's name is the same as the one being substituted for.

However, in RSLC, as shown in {\sc Subst-Lam-Capture}, we switch to a second substitution mode, {\sc Capture}, in which substituted expressions are appended to a variable's $\rep{e}$ list (in the {\sc Capture-Var-X} rule).
Similarly, when substituting for $x$ before switching to {\sc Capture} mode (in the {\sc Subst-Var-X} rule), we nondeterministically select an expression from $x$'s $\rep{e}$ list to replace $x$ with (with uniform randomness).

@@

\begin{figure}[t]
	\fbox{$\subst{e_0}{x}{e}$}
	\TypeRules{
		\SubstLamCapture \qquad \SubstLam \\
		\SubstVarX \qquad \SubstVarY \\
	}
	\fbox{$\capture{e_0}{x}{e}$}
	\TypeRules{
		\CaptureVarX \qquad \CaptureVarY \\
	}
	\caption{RSLC substitution rules. The ``$e_1 \in ...$'' premise in {\sc Subst-Var-X} represents nondeterministic choice. (Only important rules are shown here; Figure~\ref{fig:subst-complete} shows the complete set.)}
	\label{fig:subst}
\end{figure}

\subsection{Evaluation Example}

As an example, I present the steps for evaluating the $(\lambda x. (\lambda x. x))~A~B$ from the introduction:

\begin{eqnarray*}
	& & (\lambda x. (\lambda x. x[]))~A~B \\
	& \mapsto & (\subst{A}{x} (\lambda x. x[])) B \\
	& = & (\lambda x. \capture{A}{x} x[]) B \\
	& = & (\lambda x. x[A]) B \\
	& \mapsto & \subst{B}{x} x[A] \\
	& = & \text{\em (any element of the list $[B,A]$)}
\end{eqnarray*}

\section{Random Programming with RSLC}

Already we have a system with which we can implement simple random number generators, such as a simple 6-sided die:
\[
(\lambda x.
(\lambda x.
(\lambda x.
(\lambda x.
(\lambda x.
(\lambda x. x))))))
~\epsdice{1}
~\epsdice{2}
~\epsdice{3}
~\epsdice{4}
~\epsdice{5}
~\epsdice{6}
\]

However,

\newpage

\section{Evaluation}

We computed some random numbers woo

\begin{figure}[t]
	\begin{center}
	\includegraphics[width=0.48\textwidth]{rslc-rdr.pdf}
	\end{center}
	\caption{We show that RSLC is a generalization of the Random Distance Run~\cite{rdr}.}
	\label{fig:rdr}
\end{figure}

\begin{figure}[t]
	\begin{center}
	\includegraphics[width=0.40\textwidth]{rslc-d20.pdf}
	\end{center}
	\caption{RSLC can also be used to implement popular role-playing games.}
	\label{fig:rdr}
\end{figure}

@@

\section{Discussion of Substitution Rules}
\begin{figure*}[t]
	\fbox{$\subst{e_0}{x}{e}$}
	\TypeRules{
		\SubstLamCapture \qquad \SubstLam \qquad
		\SubstApp \\
		\SubstVarX \qquad \SubstVarY \\
		\SubstCaseCapture \\ \SubstCase \\
		\SubstZero \qquad \SubstSuc \\
		\SubstLetCapture \qquad \SubstLet
	}
	\fbox{$\capture{e_0}{x}{e}$}
	\TypeRules{
		\CaptureLam \qquad \CaptureApp \qquad
		\CaptureVarX \qquad \CaptureVarY \\
		\CaptureCase \\
		\CaptureZero \qquad \CaptureSuc \qquad \CaptureLet
	}
	\caption{Complete RSLC substitution and capture rules.}
	\label{fig:subst-complete}
\end{figure*}
